---
title: "Homework 1"
author: "Hithu Oleti"
output: html_document
date: "2025-02-04"
---
## Question 1:
### Adding Header Rows to the Data

```{r}
Typical_Employee_Survey_Data <- read.csv("Typical_Employee_Survey_Data.csv")

colnames(Typical_Employee_Survey_Data) <- c("Age", "Gender", "Job Satisfaction", "ImportantJobCharacteristic", "YearsWorked", "PromotionLikelihood", "WorkDecisionPower", "BudgetDecParticipation", "ProudWorker", "TurnDownJob", "RelationsScale")
print(Typical_Employee_Survey_Data)

```
## Question 2.1: 
### 5-point Summary and Boxplot for Age 

```{r}
fivenum(Typical_Employee_Survey_Data$Age)
boxplot(Typical_Employee_Survey_Data$Age, main = "Distribution of Employee Ages", ylab = "Age") 
```

## Question 2.2
### 5-point Summary and Boxplot for Years Worked 
```{r}
fivenum(Typical_Employee_Survey_Data$YearsWorked)
boxplot(Typical_Employee_Survey_Data$YearsWorked, main = "Distribution of Amount of Years Employees Worked", ylab = "Years Worked")

```

## Question 3.1: 
### Convert Gender to Factors 
```{r}
Typical_Employee_Survey_Data$Gender <- factor(Typical_Employee_Survey_Data$Gender, levels = c(1, 2),labels = c("Male", "Female"))


print(Typical_Employee_Survey_Data$Gender)

```
## Question 3.2: 
### Convert Job Satisfaction to Factors

```{r}
Typical_Employee_Survey_Data[,"Job Satisfaction"] <- factor(Typical_Employee_Survey_Data[,"Job Satisfaction"],levels = c(1, 2, 3, 4), labels = c("Very Satisified", "Moderately Satisfied", "A Little Dissatisfied", "Very Dissatisified"))

print(Typical_Employee_Survey_Data[,"Job Satisfaction"])


```
## Question 3.3: 
### Convert Important Job Characteristic to Factors

```{r}
Typical_Employee_Survey_Data$ImportantJobCharacteristic <- factor(Typical_Employee_Survey_Data$ImportantJobCharacteristic, levels = c(1, 2, 3, 4, 5),labels = c("High Income", "No Danger of Being Fired", "Flexible Hours", "Opportunities for Advancement", "Enjoying the Work" ))


print(Typical_Employee_Survey_Data$ImportantJobCharacteristic)

```
## Question 3.4: 
### Convert Promotion Likelihood to Factors

```{r}
Typical_Employee_Survey_Data$PromotionLikelihood <- factor(Typical_Employee_Survey_Data$PromotionLikelihood, levels = c(1, 2, 3, 4, 5),labels = c("Very Likely", "Likely", "Not Sure", "Unlikely", "Very Unlikely" ))

print(Typical_Employee_Survey_Data$PromotionLikelihood)

```

## Question 3.5: 
### Convert Work Decision Power to Factors

```{r}
Typical_Employee_Survey_Data$WorkDecisionPower <- factor(Typical_Employee_Survey_Data$WorkDecisionPower, levels = c(1, 2, 3, 4),labels = c("Always", "Much of the Time", "Sometimes", "Never" ))

print(Typical_Employee_Survey_Data$WorkDecisionPower)

```

## Question 3.6: 
### Convert Budgetary Decision Participation to Factors

```{r}
Typical_Employee_Survey_Data$BudgetDecParticipation <- factor(Typical_Employee_Survey_Data$BudgetDecParticipation, levels = c(1, 2),labels = c("Yes", "No" ))

print(Typical_Employee_Survey_Data$BudgetDecParticipation)

```
## Question 3.7: 
### Convert Proud Worker Status to Factors

```{r}
Typical_Employee_Survey_Data$ProudWorker <- factor(Typical_Employee_Survey_Data$ProudWorker, levels = c(1, 2, 3, 4),labels = c("Very Proud", "Somewhat Proud", "Indifferent", "Not At All Proud" ))

print(Typical_Employee_Survey_Data$ProudWorker)

```

## Question 3.8: 
### Convert Turning Down a Job for Higher Pay to Factors

```{r}
Typical_Employee_Survey_Data$TurnDownJob <- factor(Typical_Employee_Survey_Data$TurnDownJob, levels = c(1, 2, 3, 4, 5),labels = c("Very Likely", "Likely", "Not Sure", "Unlikely", "Very Unlikely" ))

print(Typical_Employee_Survey_Data$TurnDownJob)

```
## Question 3.9: 
### Convert Relations Scale to Factors

```{r}
Typical_Employee_Survey_Data$RelationsScale <- factor(Typical_Employee_Survey_Data$RelationsScale, levels = c(1, 2, 3, 4, 5),labels = c("Very Good", "Good", "So So", "Bad", "Very Bad" ))

print(Typical_Employee_Survey_Data$RelationsScale)

```
## Revised Whole Table
```{r}
print(Typical_Employee_Survey_Data)

```
## Question 4
### Appropriate Plots for Each Column

```{r}
par(mar=c(5,4,4,2))
gender_counts <- table(Typical_Employee_Survey_Data$Gender)
pie(gender_counts, main = "Gender Proportions", col = c("blue", "pink"))
mtext("The company is majority male.", side = 1, line = 2, cex = 1.2)

satisfaction_counts <- table(Typical_Employee_Survey_Data[,"Job Satisfaction"])
barplot(satisfaction_counts, main = "Job Satisfaction", col = "Blue")
mtext("The majority of the company feels very satisfied with their job.", side = 1, line = 3, cex = 1.2)


par(mar = c(5, 8, 4, 8))  # Adjust left (8) and right (8) margins
jobcharacteristic_counts <-table(Typical_Employee_Survey_Data$ImportantJobCharacteristic)
pie(jobcharacteristic_counts, main = "Employees' Most Important Job Characteristic")
mtext("The majority of the company feels that the most important characteristic of their job is enjoying the work.", side = 1, line = 2, cex = 0.8)


hist(Typical_Employee_Survey_Data$YearsWorked, main = "Distribution of Years Worked", xlab = "Years Worked", col = "Red")
mtext("Most people who work at the company have worked there for less than 10 years.", side = 1, line = 4, cex = 1.2)

promotionlikelihood_counts <- table(Typical_Employee_Survey_Data$PromotionLikelihood)
barplot(promotionlikelihood_counts, main = "The Likelihood of Promotion Amongst Employees", xlab = "Likelihood", ylab = "Frequency", col = "Yellow")
mtext("More people believe they are very unlikely to get promoted.", side = 1, line = 4, cex = 1.2)

workdecisionpower_count <- table(Typical_Employee_Survey_Data$WorkDecisionPower)
barplot(workdecisionpower_count, main = "Does Your Job Let You Have Power Over Work Decisions", xlab = "Frequency of Power", ylab = "Frequency", col = "Blue")
mtext("More people believe they get work decision power much of the time.", side = 1, line = 4, cex = 1.2)

budgetarydecision_count <- table(Typical_Employee_Survey_Data$BudgetDecParticipation)
pie(budgetarydecision_count, main = "Do You Participate in Budgetary Decisions?", col = c("green", "red"))
mtext("A small majority of the company believes they do participate in budgetary decisions.", side = 1, line = 2, cex = 1)

proud_count <- table(Typical_Employee_Survey_Data$ProudWorker)
barplot(proud_count, main = "How Proud Are You To Work Here?", xlab = "Proud Level", ylab = "Frequency", col = "Red")
mtext("Most people are somewhat proud to work at this company.", side = 1, line = 4, cex = 1.2)

turndownjob_count <- table(Typical_Employee_Survey_Data$TurnDownJob)
barplot(turndownjob_count, main = "How Likely Are You To Turn Down a Job With More Pay?", xlab = "Likelihood", ylab = "Frequency", col = "Yellow")
mtext("Employees are unlikely to turn down another job offer with more pay.", side = 1, line = 4, cex = 1.2)


relations_count <- table(Typical_Employee_Survey_Data$RelationsScale)
pie(relations_count, main = "How Would You Describe Relations in Your Workplace?")
mtext("Employees believe relations between coworkers are very good.", side = 1, line = 2, cex = 1.2)


```
## Question 5
### Scatter Plots for Few Relationships

```{r}
plot(Typical_Employee_Survey_Data$Age, Typical_Employee_Survey_Data$YearsWorked, main = "Age vs. Years Worked", xlab = "Age", ylab = "Years Worked", col = "Pink") 
mtext("The older the employee is, the more years they have worked at the company.", side = 1, line = 4, cex = 1.2)


plot(as.numeric(Typical_Employee_Survey_Data$TurnDownJob), Typical_Employee_Survey_Data$Age, main = "Likelihood of Promotion vs. Age", xlab = "Promotion Likelihood (1 being very likely and 5 being very unlikely", ylab = "Age", col = "Purple")
mtext("Most employees, despite their age, are unlikely to turn down a job with more pay.", side = 1, line = 4, cex = 1)



plot(as.numeric(Typical_Employee_Survey_Data$WorkDecisionPower), Typical_Employee_Survey_Data$YearsWorked, main = "Power over Work Decisions vs. Years Worked", xlab = "Power over Work Decisions (1 being always and 4 being never)", ylab = "Years Worked", col = "Blue")
mtext("There are no extremeties in this data. Employees participate in decisions sometimes/much of the time.", side = 1, line = 4, cex = 0.8)


```


